---
title: "Midterm"
author: "Yetsy Santa"
date: "03-14-2024"
output: html_document
---


library(ggplot2)
library(tidyverse)
library(viridis)  

brfss22[1:10,1:6]
attach(brfss22)
summary(brfss22)

---  
Question 1:

  Questions: Is there relationship between marijuana usage with Age?  
             Is there a statistical significant difference in the number of marijuana usage days between those who are at age 32 or age 52?
  
  Level of significance: 0.05

  Null hypothesis: There is no relationship between marijuana usage and Age.
  
  standard deviation: N=13 , mean= 
    df=12
  Standard Error: 
  
  95% confidence interval
  p= 2.179
  
```{r}
{r eval=FALSE}
   xtabs(~ use_marijan + Age_midpt)  
```


brfss22$Age_midpt <- fct_recode(brfss22$X_AGEG5YR, "21" = "Age 18 to 24",
                                "27" = "Age 25 to 29", "32" = "Age 30 to 34",
                                "37" = "Age 35 to 39", "42" = "Age 40 to 44",
                                "47" = "Age 45 to 49", "52" = "Age 50 to 54",
                                "57" = "Age 55 to 59", "62" = "Age 60 to 64",
                                "67" = "Age 65 to 69", "72" = "Age 70 to 74",
                                "77" = "Age 75 to 79", "82" = "Age 80 or older",
                                NULL = "Dont know/Refused/Missing")
brfss22$Age_midpt <- as.numeric(levels(brfss22$Age_midpt))[brfss22$Age_midpt]  

brfss22$use_marijan <- fct_recode(brfss22$marijan1, 
                                  "0" = "zero days", 
                                  "1" = "use more than one day")
brfss22$use_marijan <- as.numeric(levels(brfss22$use_marijan))[brfss22$use_marijan]



---

  Question 2:

  Questions: 
              Is there relationship between marijuana usage with educational                 level?  
              
              Is there a statistical significant difference in marijuana usage               between those who reported "Grades 9 through 11 (Some high                     school)" and those who reported "College 1 year to 3 years (Some               college or technical school)"?

Null hypothesis: There is no relationship between marijuana usage and educational level.

 
```{r}
{r eval=FALSE}
educ_shortlabel <- fct_recode(EDUCA, "never" = "Never attended school or only kindergarten",
                    "elementary" = "Grades 1 through 8 (Elementary)", 
                    "some hs" = "Grades 9 through 11 (Some high school)",
                    "hs" = "Grade 12 or GED (High school graduate)",
                    "some coll" = "College 1 year to 3 years (Some college or technical school)",
                    "degree" = "College 4 years or more (College graduate)",
                    "Refused" = "Refused")

xtabs(~ use_marijan + educ_shortlabel)
```
   
``` {r}
library(stats)
```

``` {r}
sub_data_some_hs <- subset(brfss22, EDUCA == "Grades 9 through 11 (Some high school)" & !is.na(MARIJAN1))
sub_data_some_coll <- subset(brfss22, EDUCA == "College 1 year to 3 years (Some college or technical school)" & !is.na(MARIJAN1))
```

``` {r}
anova_result <- aov(MARIJAN1 ~ EDUCA, data = brfss22)
summary(anova_result)
```
The data shows that p< 2e-16 which is less than the level of significance of 5%. Therefore, we reject the null hypothesis. This means that there is a significant difference marijuana usage with educational levels.
  
---
Question 3:

How marijuana use correlates with education?

```{r}
NN <- length(brfss22$EDUCA)
set.seed(12345)
restrict_1 <- (runif(NN) < 0.1) # use just 10% 
summary(restrict_1)
```

```{r}
brfss_small <- subset(brfss22, restrict_1)
```



```{r}
ggplot(brfss_small, aes(x = MARIJAN1, fill = EDUCA)) + geom_bar()
```

```{r}
ggplot(brfss_small, aes(x = EDUCA, fill = MARIJAN1)) + geom_bar()
```
  



  
  
```{r}
brfss_small_hs <- subset(brfss_small,brfss_small$EDUCA == "Grades 9 through 11 (Some high school)")
brfss_small_coll <- subset(brfss_small,brfss_small$EDUCA == "College 1 year to 3 years (Some college or technical school)")
```

```{r}
ggplot(brfss_small_hs, aes(x = MARIJAN1, fill = EDUCA)) + geom_bar(position = "fill") + scale_fill_viridis_d(option="rocket")
```

  
X1 <- standardize_varb_to01(brfss22$EDUCA)
X2 <- standardize_varb_to01(brfss22$MARIJAN)
X3 <- brfss22$EDUCA_recode
X4 <- brfss22$MARIJAN_recode
# you could add more X variables...
Y <- brfss22$MARIJAN_recode









